---
output:
  html_document:
    keep_md: yes
---
# Reproducible Research: Peer Assessment 1
Author: Kriss


## Introduction
This is the first peer assessment of Reproducible Research of JH on coursera. This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.


### Loading and preprocessing the data

1. Load the data.
```{r load}
data <- read.csv(file = "activity.csv", sep = ",", header = TRUE)
```

2. Process the data.
```{r process}
no.na <- na.omit(data)
```


### What is mean total number of steps taken per day?

1. Make a histogram of the total number of steps taken each day.
```{r hist}
dailysteps <- data.frame(date = format(as.Date(names(tapply(no.na$steps, no.na$date, sum))), "%m%d"), steps = tapply(no.na$steps, no.na$date, sum))
library(lattice)
barchart(steps ~ date, data = dailysteps, xlab = "day", main = "Daily Steps from October 1 to November 30, 2012", horizontal = FALSE, scales = list(x = list(cex = 0.7)), xlim = c(0, 62), ylim = c(0, 25000))
```

2. Calculate and report the mean and median total number of steps taken per day

Here is the code to calculate the mean and median:
```{r m,results='hide'}
meansteps <- mean(dailysteps$steps, na.rm = TRUE)
mediansteps <- median(dailysteps$steps, na.rm = TRUE)
```
The result is: `r meansteps` and `r mediansteps`


### What is the average daily activity pattern?

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r average}
averagesteps <- data.frame(interval = names(tapply(no.na$steps, no.na$interval, mean)), average = tapply(no.na$steps, no.na$interval, mean))
levels(averagesteps$interval) <- order(unique(no.na$interval))
averagesteps$interval <- sort(averagesteps$interval)
xyplot(average ~ interval, data = averagesteps, type = "b", xlim = c(0, 300), ylim = c(0, 250), main = "Average across all the days")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r max,results='hide'}
maxaverage <- which.max(averagesteps$average)
```

The maximum number of steps occur at `r maxaverage`e 5-minute interval.


### Imputing missing values
Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r row,results='hide'}
narows <- nrow(data) - nrow(no.na)
```

The total number of missing values in the dataset: `r narows` rows.


2. Fill the means of 5-minute interval accordingly in all of the missing values in the dataset.
```{r new}
a <- aggregate(steps ~ interval, data, mean, na.action = na.omit)
m <- merge(data, a, by.x = "interval", by.y = "interval")
m <- m[order(m$date),]
nas <- subset(m, is.na(m$steps.x))
m$steps.x[is.na(m$steps.x)] <- nas$steps.y
newdata <- data.frame(steps = m$steps.x, date = m$date, interval = m$interval)
```
3. Make a histogram of the total number of steps taken each day with the new dataset.
```{r newhist}
newsteps <- aggregate(steps ~ date, newdata, sum)
barchart(steps ~ date, data = newsteps, xlab = "day", main = "Daily Steps from October 1 to November 30, 2012", xlim = c(0, 62), ylim = c(0, 25000))
```

4. Calculate and report the mean and median total number of steps taken per day with the new dataset.
```{r newm}
newmean <- mean(newsteps$steps)
newmedian <- median(newsteps$steps)
```
The result is: `r newmean` and `r newmedian`

The mean values are same, but the median values have 1 step diffrence between the new and old dataset. After imputing missing data, it is easier to judge everyday steps by the plot.


### Are there differences in activity patterns between weekdays and weekends?

1. With the new dataset, create a new factor variable with two levels  indicating whether a given date is a weekday or weekend day.
```{r week}
Sys.setlocale("LC_TIME", "English")
newdata$weekdays <- weekdays(as.Date(newdata$date))
weekday <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
weekend <- c("Saturday", "Sunday")
newdata[which(newdata$weekdays %in% weekday), 4] <- "weekday"
newdata[which(newdata$weekdays %in% weekend), 4] <- "weekend"
```

2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r weekpattern}
weekdayaverage <- aggregate(steps ~ interval, newdata[which(newdata$weekdays == "weekday"), ], mean)
weekendaverage <- aggregate(steps ~ interval, newdata[which(newdata$weekdays == "weekend"), ], mean)
wda <- cbind(weekdayaverage, weekdays = rep("weekday", length(weekdayaverage$steps)))
wea <- cbind(weekendaverage, weekdays = rep("weekend", length(weekendaverage$steps)))
weekaverage <- rbind(wda,wea)
xyplot(steps ~ interval | weekdays, data = weekaverage, type = "l", ylab = "Number of steps", main = "Activity patterns of weekdays and weekends", layout = c(1, 2))
```

